---
title: "Lab 03 - Nobel laureates"
author: "Ben Waggener"
date: "01/31/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

"How many observations and how many variables are in the dataset? Use inline code to answer this question. What does each row represent?"

```{r read-data}
names(nobel)
```

```{r summaryofdata}
glimpse(nobel)

```


There are 26 variables and 935 observations. Each row represents an observation (a nobel laureate).



### Exercise 2

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...
